Simulating High-Dimensional Multivariate Data

### 2021-02-24 020-simulation-algorithm.Rmd
Add new section 3.3

## `rvec` for random vector generation {#rand-vec-gen}

- Discuss how the HD NORTA defective problem and the scalability issues lead to our algorithm
We now describe `rvec`, our algorithm to generate random vectors, which mirrors the classical NORTA algorithm above with some modifications for rank-based dependency matching:

@@

>>>>

## Random vector generator {#rand-vec-gen}

We now describe our algorithm to generate random vectors, which mirrors the classical NORTA algorithm above with some modifications for rank-based dependency matching:

@@

## Algorithm details

*Correlation Conversion*

Given:

* A target Spearman's $\rho$ or Kendall's $\tau$ correlation

Do:

* Determine the Pearson correlation for use in a Gaussian copula such that the resulting random samples using the NORTA algorithm have an estimated Spearman or Kendall correlation that matches the target correlation

From Proposition 22 (Lebrun and Dutfow, 2009), we have the result for Gaussian copula that the copula correlation $r$ is related to Spearman's $\rho$ and Kendall's $\tau$ by the following relations:

$r = 2 \sin \left(\frac{\pi}{6} \rho\right) \\
r = \sin \left(\frac{\pi}{2} \tau\right)$

* Lebrun, R., & Dutfoy, A. (2009). An innovating analysis of the Nataf transformation from the copula viewpoint. Probabilistic Engineering Mechanics, 24(3), 312-320.

*Pearson Matching*

Given:

* A target Pearson correlation matrix, $R$
* A list of marginal distributions, $F$

Do:

* Determine the input correlation, $\tilde{P}$, such that the resulting random samples using the NORTA algorithm have an estimated Pearson correlation that matches $P$

This method relies on Equations 19, 39, and 49 (Xiao and Zhou, 2019).

* Xiao, Q., & Zhou, S. (2019). Matching a correlation coefficient by a Gaussian copula. Communications in Statistics-Theory and Methods, 48(7), 1728-1747.

*Nearest Correlation Matrix*

This algorithm is trying to find a correlation matrix $X$ close to $G$ in the convex optimization problem:

$\begin{aligned}
    \mathrm{min}\quad & \frac{1}{2} \Vert G - X \Vert^2 \\
    \mathrm{s.t.}\quad & X_{ii} = 1, \quad i = 1, \ldots , n, \\
    & X \in S_{+}^{n}
\end{aligned}$

* Qi, H., & Sun, D. (2006). A quadratically convergent Newton method for computing the nearest correlation matrix. SIAM journal on matrix analysis and applications, 28(2), 360-385.
